---
title: "Fake News Epidemics and Detection"
author: "Alejandro Tejada & Eloy Marín"
date: "15/01/2022"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show

  pdf_document: default
urlcolor: blue
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F, fig.align = "center")
```

```{r echo=FALSE, include=FALSE}
library(dplyr)
library(igraph)
library("poweRlaw")
library("ggplot2")
source("simulation_helpers.R")
source("generate_immune_graphs.R")
```

# **Fake News Epidemics and Detection**


We will analyze two different networks, FNN and TNN. FNN stands for
Fake News Network, and contains nodes that have interacted with a fake new and 
their followers as suspects to get infected. On the contrary, TNN stands for
True News Network, and contains nodes that have interacted with a true new and 
their followers as suspects to get infected.
```{r}

# Initialize fake news network
FNN <- read.graph("graph_politifact15014_fake.txt", format="edgelist", directed=FALSE)
FNN.infected <- read.table(file="infected_politifact15014_fake.txt")$V1 + 1

# Initialize true news network
TNN <- read.graph("graph_politifact1084_real.txt", format="edgelist", directed=FALSE)
TNN.infected <- read.table(file="infected_politifact1084_real.txt")$V1 + 1

```


## Metrics

We will analyze & compare these networks according to different metrics.


### Degree distribution

First, we plot a summary of the degree distribution of the two networks. We can 
see that although we have less than a half nodes for the FNN, the average degree its 
almost the double of the one for TNN. This probably indicates that FNN has a higher 
clustering coefficient, meaning that the nodes that spread fake news usually belong 
to a community or group of communities.


```{r}
degree_summary <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(degree_summary) <- c('Network', 'N', 'Max. degree', 'M/N', 'N/M')


degrees <- degree(FNN)
FNN.degree.histogram <- as.data.frame(table(degrees))
FNN.degree.histogram[,1] <- as.numeric( paste(FNN.degree.histogram[,1]))
network_degree_summary <- c('FNN',vcount(FNN),
                   max(FNN.degree.histogram$degrees),
                   sum(FNN.degree.histogram$degrees)/vcount(FNN),
                   vcount(FNN)/sum(FNN.degree.histogram$degrees))
degree_summary[nrow(degree_summary) + 1,] = network_degree_summary

degrees <- degree(TNN)
TNN.degree.histogram <- as.data.frame(table(degrees))
TNN.degree.histogram[,1] <- as.numeric( paste(TNN.degree.histogram[,1]))
network_degree_summary <- c('TNN',vcount(TNN),
                   max(TNN.degree.histogram$degrees),
                   sum(TNN.degree.histogram$degrees)/vcount(TNN),
                   vcount(TNN)/sum(TNN.degree.histogram$degrees))
degree_summary[nrow(degree_summary) + 1,] = network_degree_summary

degree_summary
```


If we check the actual degree distribution that the networks follow, we can see that
they follow a kind of truncated power-law, meaning that there are some nodes that 
can be considered as hubs, since they are connecting big amounts of nodes compared to another nodes.
This "hub" nodes may be the most interesting to remove if we want to use immunization techniques to prevent the spread of fake news.
```{r}
degrees_df <- data.frame()
degrees_df <- rbind(degrees_df, cbind(FNN.degree.histogram,
                                        group='FNN'))
degrees_df <- rbind(degrees_df, cbind(TNN.degree.histogram,
                                        group='TNN'))

ggplot(degrees_df, aes(x = degrees, y = Freq, color=as.factor(group))) +
    geom_point() +
    scale_x_continuous("Degree",
                       breaks = c(1, 3, 10, 30, 100, 300),
                       trans = "log10") +
    scale_y_continuous("Nodes",
                       breaks = c(1, 3, 10, 30, 100, 300, 1000),
                       trans = "log10") +
    ggtitle("Degree Distribution (log-log)") +
    theme_bw()
```


#### Spreader nodes degree

We want to check the degree of the nodes that start spreading news, and check if
these nodes have high degree.
```{r}
FNN.infected.n <- length(FNN.infected)
FNN.degrees <- degree(FNN)
FNN.idx.nodes.degree <- sort(FNN.degrees, decreasing = T)[1:FNN.infected.n]
FNN.nodes.degree <- which(degree(FNN) %in% FNN.idx.nodes.degree)
FNN.nodes.degree.infected <- intersect(FNN.nodes.degree, FNN.infected)

TNN.infected.n <- length(TNN.infected)
TNN.degrees <- degree(TNN)
TNN.idx.nodes.degree <- sort(TNN.degrees, decreasing = T)[1:TNN.infected.n]
TNN.nodes.degree <- which(degree(TNN) %in% TNN.idx.nodes.degree)
TNN.nodes.degree.infected <- intersect(TNN.nodes.degree, TNN.infected)
```
The `r 100*length(FNN.nodes.degree.infected)/FNN.infected.n`% of the nodes with higher degree are precisely the ones that start propagating the fake news. In the case of true news, `r 100*length(TNN.nodes.degree.infected)/TNN.infected.n`%
with higher degree are the ones that start propagating the news. Actually, most of the nodes that start propagation the news have extremely large degree compared with the average degree.


### Centrality
We will also compare both graphs w.r.t their betweenness centrality. To do so,
need to compute it first for both networks.

```{r}
FNN.betweenness <- betweenness(FNN)
TNN.betweenness <- betweenness(TNN)
```

Now we can perform the same analysis as before, checking that the nodes that
start spreading news are the ones with higher betweenness. 
```{r}
FNN.idx.nodes.betweenness <- sort(FNN.betweenness, decreasing = T)[1:FNN.infected.n]
FNN.nodes.betweenness <- which(FNN.betweenness %in% FNN.idx.nodes.betweenness)
FNN.nodes.betweenness.infected <- intersect(FNN.nodes.betweenness, FNN.infected)

TNN.idx.nodes.betweenness <- sort(TNN.betweenness, decreasing = T)[1:TNN.infected.n]
TNN.nodes.betweenness <- which(TNN.betweenness %in% TNN.idx.nodes.betweenness)
TNN.nodes.betweenness.infected <- intersect(TNN.nodes.betweenness, TNN.infected)
```
The `r 100*length(FNN.nodes.betweenness.infected)/FNN.infected.n`% of the nodes with higher betweennes are the ones that start propagating the fake news. In the case of true news, `r 100*length(TNN.nodes.betweenness.infected)/TNN.infected.n`%
with higher betweenness are the ones that start propagating the news. Actually, most of the nodes that start propagation the news have extremely large betweenness.



### Clustering

If we compare clustering coefficients of FNN and TNN, we can see that the network corresponding to fake news has larger clustering coefficients, indicating again that users that spread fake news belong to a community or group of communities.
```{r}
clustering_summary <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(clustering_summary) <- c('Network', 'Global Transitivity', 'Local Avg. Transitivity')

network_clustering_summary <- c('FNN',
                              transitivity(FNN),
                              transitivity(FNN, "localaverage"))
clustering_summary[nrow(clustering_summary) + 1,] = network_clustering_summary

network_clustering_summary <- c('TNN',
                              transitivity(TNN),
                              transitivity(TNN, "localaverage"))
clustering_summary[nrow(clustering_summary) + 1,] = network_clustering_summary

clustering_summary
```



### Examine initial infected graph

Now we will analyze the initial graph that only contains nodes that spread the 
news.
```{r}
FNN_only_infected <- subgraph(FNN, FNN.infected)
Isolated = which(degree(FNN_only_infected)==0)
FNN_only_infected = delete.vertices(FNN_only_infected, Isolated)

TNN_only_infected <- subgraph(TNN, TNN.infected)
Isolated = which(degree(TNN_only_infected)==0)
TNN_only_infected = delete.vertices(TNN_only_infected, Isolated)
```

```{r}
infected_summary <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(infected_summary) <- c('Network', 'N', 'Max. degree', 'M/N', 'N/M',
                                'Global Transitivity', 
                                'Local Avg. Transitivity', 
                                'Diameter', 'Density')

degrees <- degree(FNN_only_infected)
FNN_only_infected.degree.histogram <- as.data.frame(table(degrees))
FNN_only_infected.degree.histogram[,1] <- as.numeric(paste(FNN_only_infected.degree.histogram[,1]))
network_infected_summary <- c('FNN',vcount(FNN_only_infected),
                   max(FNN_only_infected.degree.histogram$degrees),
                   sum(FNN_only_infected.degree.histogram$degrees)/vcount(FNN_only_infected),
                   vcount(FNN_only_infected)/sum(FNN_only_infected.degree.histogram$degrees),
                   transitivity(FNN_only_infected),
                   transitivity(FNN_only_infected, "localaverage"),
                   diameter(FNN_only_infected),
                   edge_density(FNN_only_infected)
                   )
infected_summary[nrow(infected_summary) + 1,] = network_infected_summary

degrees <- degree(TNN_only_infected)
TNN_only_infected.degree.histogram <- as.data.frame(table(degrees))
TNN_only_infected.degree.histogram[,1] <- as.numeric(paste(TNN_only_infected.degree.histogram[,1]))
network_infected_summary <- c('TNN',vcount(TNN_only_infected),
                   max(TNN_only_infected.degree.histogram$degrees),
                   sum(TNN_only_infected.degree.histogram$degrees)/vcount(TNN_only_infected),
                   vcount(TNN_only_infected)/sum(TNN_only_infected.degree.histogram$degrees),
                   transitivity(TNN_only_infected),
                   transitivity(TNN_only_infected, "localaverage"),
                   diameter(TNN_only_infected),
                   edge_density(TNN_only_infected)
                   )
infected_summary[nrow(infected_summary) + 1,] = network_infected_summary

infected_summary
```

```{r}
Monte_Carlo.S <- function(value, threshold) {
  if (value > threshold) {
    return(1) }
  else {
    return(0)
  }
}

switching.transitivity <- function(graph){
    
    threshold <- transitivity(graph)
    
    edgelist_graph <- as_edgelist(graph, names = FALSE)
    
    E <- nrow(edgelist_graph)
    Q = 1
    
    times = Q*E
    success <- 0
    fails <- 0
    
    transivity_switching <- c()
    
    for (i in 1:times) {
      pair1 <- sample(1:E,1,replace=TRUE)
      pair2 <- sample(1:E,1,replace=TRUE)
      
      node1 <- edgelist_graph[pair1, 2]
      node2 <- edgelist_graph[pair2, 2]
      
      if ((node1 != edgelist_graph[pair2, 1]) & #avoid loops 
          (node2 != edgelist_graph[pair1, 1])) {
        provisional_edge_list <- edgelist_graph
        provisional_edge_list[pair1,2] <- node2
        provisional_edge_list[pair2,2] <- node1  
        new_E <- nrow(unique(provisional_edge_list))
      } else {
        fails <- fails + 1
        return
      }
      
      if (E == new_E) { # if the new pair created doesn't exist before
        edgelist_graph <- provisional_edge_list
        success <- success + 1
      } else {
        fails <- fails + 1
        return
      }
      
      new_file_graph <- graph_from_edgelist(edgelist_graph)
      transivity_switching <- c(transivity_switching, transitivity(new_file_graph))
      
    }

    greater_vector <- sapply(transivity_switching, Monte_Carlo.S, threshold=threshold)
    
  return(greater_vector)
}


switching.transitivity.localaverage <- function(graph){
    
    threshold <- transitivity(graph, "localaverage")
    
    edgelist_graph <- as_edgelist(graph, names = FALSE)
    
    E <- nrow(edgelist_graph)
    Q = 1
    
    times = Q*E
    success <- 0
    fails <- 0
    
    transivity_switching <- c()
    
    for (i in 1:times) {
      pair1 <- sample(1:E,1,replace=TRUE)
      pair2 <- sample(1:E,1,replace=TRUE)
      
      node1 <- edgelist_graph[pair1, 2]
      node2 <- edgelist_graph[pair2, 2]
      
      if ((node1 != edgelist_graph[pair2, 1]) & #avoid loops 
          (node2 != edgelist_graph[pair1, 1])) {
        provisional_edge_list <- edgelist_graph
        provisional_edge_list[pair1,2] <- node2
        provisional_edge_list[pair2,2] <- node1  
        new_E <- nrow(unique(provisional_edge_list))
      } else {
        fails <- fails + 1
        return
      }
      
      if (E == new_E) { # if the new pair created doesn't exist before
        edgelist_graph <- provisional_edge_list
        success <- success + 1
      } else {
        fails <- fails + 1
        return
      }
      
      new_file_graph <- graph_from_edgelist(edgelist_graph)
      transivity_switching <- c(transivity_switching, transitivity(new_file_graph, "localaverage"))
      
    }

    greater_vector <- sapply(transivity_switching, Monte_Carlo.S, threshold=threshold)
    
  return(greater_vector)
}

switching.edge.density <- function(graph){
    
    threshold <- edge_density(graph)
    
    edgelist_graph <- as_edgelist(graph, names = FALSE)
    
    E <- nrow(edgelist_graph)
    Q = 1
    
    times = Q*E
    success <- 0
    fails <- 0
    
    transivity_switching <- c()
    
    for (i in 1:times) {
      pair1 <- sample(1:E,1,replace=TRUE)
      pair2 <- sample(1:E,1,replace=TRUE)
      
      node1 <- edgelist_graph[pair1, 2]
      node2 <- edgelist_graph[pair2, 2]
      
      if ((node1 != edgelist_graph[pair2, 1]) & #avoid loops 
          (node2 != edgelist_graph[pair1, 1])) {
        provisional_edge_list <- edgelist_graph
        provisional_edge_list[pair1,2] <- node2
        provisional_edge_list[pair2,2] <- node1  
        new_E <- nrow(unique(provisional_edge_list))
      } else {
        fails <- fails + 1
        return
      }
      
      if (E == new_E) { # if the new pair created doesn't exist before
        edgelist_graph <- provisional_edge_list
        success <- success + 1
      } else {
        fails <- fails + 1
        return
      }
      
      new_file_graph <- graph_from_edgelist(edgelist_graph)
      transivity_switching <- c(transivity_switching, edge_density(new_file_graph))
      
    }

    greater_vector <- sapply(transivity_switching, Monte_Carlo.S, threshold=threshold)
    
  return(greater_vector)
}
```


```{r}
sum(switching.edge.density(FNN_only_infected))
sum(switching.edge.density(TNN_only_infected))

sum(switching.transitivity(FNN_only_infected))
sum(switching.transitivity(TNN_only_infected))

sum(switching.transitivity.localaverage(FNN_only_infected))
sum(switching.transitivity.localaverage(TNN_only_infected))
```

## Epidemic simulation

After comparing metrics between networks, we will compare the spread of news in these networks following a SI epidemic model. The simulation is performed in C++, and it outputs the ratio of infected nodes at each time step. Simulations were performed with $\beta =  0.1$ and $iters=100$. 

We can observe that, in terms of epidemic modeling, there is no significant difference between FNN and TNN networks.

```{r}
FNN.infected_ratio = as.numeric(scan("si_sim_politifact15014.csv", sep=',', what = "numeric", quiet = TRUE))
TNN.infected_ratio = as.numeric(scan("si_sim_politifact1084.csv", sep=',', what = "numeric", quiet = TRUE))
simulations_df <- data.frame()

plot(x=1:100, y=FNN.infected_ratio, main="SI simulation", xlab="Time",
     ylab="Infected ratio", type="l", col="red")
lines(x=1:100,y=TNN.infected_ratio, type="l", col="green")
legend(x = "topright",          # Position
       legend = c("FNN", "TNN"),  # Legend texts
       lty = c(1, 2),           # Line types
       col = c(2, 3),           # Line colors
       lwd = 2)                 # Line width
```

## Generate immune graphs

To immunize the graphs, we propose to analyze the 1000 nodes with higher betweenness and remove them. We will only remove the ones that do not participate in the original spreading of information, and we will do it in three different steps. First we will remove just one third of the nodes, then two thirds and finally all of the 1000 nodes with high betweenness.

```{r}
graph.file <- c("graph_politifact15014_fake.txt")
infected.file <- c("infected_politifact15014_fake.txt")

generate.inmune.graphs(graph.file, infected.file)
```

```{r}
tmax <- 100

no_immune <- read.table("immune0.csv", sep = ",", nrows = 1)
immune1 <- read.table("immune1.csv", sep = ",", nrows = 1)
immune2 <- read.table("immune2.csv", sep = ",", nrows = 1)
immune3 <- read.table("immune3.csv", sep = ",", nrows = 1)

no_immune <- unname(unlist(no_immune))
immune1 <- unname(unlist(immune1))
immune2 <- unname(unlist(immune2))
immune3 <- unname(unlist(immune3))

plot(no_immune,
     xlim = c(1, tmax), ylim = c(0, 1),
     xlab = "Time", ylab = "Infection rate", main = "Original Graph"
)

plot(immune1,
     xlim = c(1, tmax), ylim = c(0, 1),
     xlab = "Time", ylab = "Infection rate", main = "Immune Graph 1"
)

plot(immune2,
     xlim = c(1, tmax), ylim = c(0, 1),
     xlab = "Time", ylab = "Infection rate", main = "Immune Graph 2"
)

plot(immune3,
     xlim = c(1, tmax), ylim = c(0, 1),
     xlab = "Time", ylab = "Infection rate", main = "Immune Graph 3"
)
```