---
title: "Fake News Epidemics and Detection"
author: "Alejandro Tejada & Eloy Marín"
date: "15/01/2022"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show

  pdf_document: default
urlcolor: blue
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F, fig.align = "center")
```

```{r echo=FALSE, include=FALSE}
library(dplyr)
library(igraph)
library("poweRlaw")
library("ggplot2")
```

# **Fake News Epidemics and Detection**

We will analyze two different networks, FNN and TNN. FNN stands for
Fake News Network, and contains nodes that have interacted with fake news and 
their followers as suspects to get infected. On the contrary, TNN stands for
True News Network, and contains nodes that have interacted with true news and 
their followers as suspects to get infected.

```{r}
# Initialize fake news network
FNN <- read.graph("graph_politifact15014_fake.txt", format="edgelist", directed=FALSE)
FNN.infected <- read.table(file="infected_politifact15014_fake.txt")$V1 + 1

# Initialize true news network
TNN <- read.graph("graph_politifact1084_real.txt", format="edgelist", directed=FALSE)
TNN.infected <- read.table(file="infected_politifact1084_real.txt")$V1 + 1

```


We will analyze & compare these networks according to different metrics.


## Degree distribution

First, we plot a summary of the degree distribution of the two networks. We can 
see that although we have less than a half nodes for the FNN, the average degree its 
almost the double of the one for TNN. This probably indicates that FNN has a higher 
clustering coefficient, meaning that the nodes that spread fake news usually belong 
to a community or group of communities.


```{r}
degree_summary <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(degree_summary) <- c('Network', 'N', 'Max. degree', 'M/N', 'N/M')


degrees <- degree(FNN)
FNN.degree.histogram <- as.data.frame(table(degrees))
FNN.degree.histogram[,1] <- as.numeric( paste(FNN.degree.histogram[,1]))
network_degree_summary <- c('FNN',vcount(FNN),
                   max(FNN.degree.histogram$degrees),
                   sum(FNN.degree.histogram$degrees)/vcount(FNN),
                   vcount(FNN)/sum(FNN.degree.histogram$degrees))
degree_summary[nrow(degree_summary) + 1,] = network_degree_summary

degrees <- degree(TNN)
TNN.degree.histogram <- as.data.frame(table(degrees))
TNN.degree.histogram[,1] <- as.numeric( paste(TNN.degree.histogram[,1]))
network_degree_summary <- c('TNN',vcount(TNN),
                   max(TNN.degree.histogram$degrees),
                   sum(TNN.degree.histogram$degrees)/vcount(TNN),
                   vcount(TNN)/sum(TNN.degree.histogram$degrees))
degree_summary[nrow(degree_summary) + 1,] = network_degree_summary

degree_summary
```


If we check the actual degree distribution that the networks follow, we can see that
they follow a kind of truncated power-law, meaning that there are some nodes that 
can be considered as hubs, since they are connecting big amounts of nodes compared to another nodes.
This "hub" nodes may be the most interesting to remove if we want to use immunization techniques to prevent
the spread of fake news.
```{r}
degrees_df <- data.frame()
degrees_df <- rbind(degrees_df, cbind(FNN.degree.histogram,
                                        group='FNN'))
degrees_df <- rbind(degrees_df, cbind(TNN.degree.histogram,
                                        group='TNN'))

ggplot(degrees_df, aes(x = degrees, y = Freq, color=as.factor(group))) +
    geom_point() +
    scale_x_continuous("Degree",
                       breaks = c(1, 3, 10, 30, 100, 300),
                       trans = "log10") +
    scale_y_continuous("Nodes",
                       breaks = c(1, 3, 10, 30, 100, 300, 1000),
                       trans = "log10") +
    ggtitle("Degree Distribution (log-log)") +
    theme_bw()
```


### Spreader nodes degree

We want to check the degree of the nodes that start spreading news, and check if
these nodes have high degree.
```{r}
FNN.infected.n <- length(FNN.infected)
FNN.degrees <- degree(FNN)
FNN.idx.nodes.degree <- sort(FNN.degrees, decreasing = T)[1:FNN.infected.n]
FNN.nodes.degree <- which(degree(FNN) %in% FNN.idx.nodes.degree)
FNN.nodes.degree.infected <- intersect(FNN.nodes.degree, FNN.infected)

TNN.infected.n <- length(TNN.infected)
TNN.degrees <- degree(TNN)
TNN.idx.nodes.degree <- sort(TNN.degrees, decreasing = T)[1:TNN.infected.n]
TNN.nodes.degree <- which(degree(TNN) %in% TNN.idx.nodes.degree)
TNN.nodes.degree.infected <- intersect(TNN.nodes.degree, TNN.infected)
```
The `r 100*length(FNN.nodes.degree.infected)/FNN.infected.n`% of the nodes with higher degree are precisely the ones that start propagating the fake news. In the case of true news, `r 100*length(TNN.nodes.degree.infected)/TNN.infected.n`%
with higher degree are the ones that start propagating the news. Actually, most of the nodes that start propagation the news have extremely large degree compared with the average degree.


## Centrality
We will also compare both graphs w.r.t their betweenness centrality. To do so,
need to compute it first.

```{r}
FNN.betweenness <- betweenness(FNN)
TNN.betweenness <- betweenness(TNN)
```

Now we can perform the same analysis as before, checking that the nodes that
start spreading news are the ones with higher betweenness. 
```{r}
FNN.idx.nodes.betweenness <- sort(FNN.betweenness, decreasing = T)[1:FNN.infected.n]
FNN.nodes.betweenness <- which(FNN.betweenness %in% FNN.idx.nodes.betweenness)
FNN.nodes.betweenness.infected <- intersect(FNN.nodes.betweenness, FNN.infected)

TNN.idx.nodes.betweenness <- sort(TNN.betweenness, decreasing = T)[1:TNN.infected.n]
TNN.nodes.betweenness <- which(TNN.betweenness %in% TNN.idx.nodes.betweenness)
TNN.nodes.betweenness.infected <- intersect(TNN.nodes.betweenness, TNN.infected)
```
The `r 100*length(FNN.nodes.betweenness.infected)/FNN.infected.n`% of the nodes with higher betweennes are the ones that start propagating the fake news. In the case of true news, `r 100*length(TNN.nodes.betweenness.infected)/TNN.infected.n`%
with higher betweennes are the ones that start propagating the news. Actually, most of the nodes that start propagation the news have extremely large betweenness.


## Examine initial infected graph

Now we will analyze the initial graph that only contains nodes that spread the 
news.
```{r}
FNN_only_infected <- subgraph(FNN, FNN.infected)
Isolated = which(degree(FNN_only_infected)==0)
FNN_only_infected = delete.vertices(FNN_only_infected, Isolated)

TNN_only_infected <- subgraph(TNN, TNN.infected)
Isolated = which(degree(TNN_only_infected)==0)
TNN_only_infected = delete.vertices(TNN_only_infected, Isolated)
```

```{r}

infected_summary <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(infected_summary) <- c('Network', 'Transitivity', 'Diameter', 'Density')

network_infected_summary <- c('FNN',
                              transitivity(FNN_only_infected, "localaverage"),
                              diameter(FNN_only_infected),
                              edge_density(FNN_only_infected))
infected_summary[nrow(infected_summary) + 1,] = network_infected_summary

network_infected_summary <- c('TNN',
                              transitivity(TNN_only_infected, "localaverage"),
                              diameter(TNN_only_infected),
                              edge_density(TNN_only_infected))
infected_summary[nrow(infected_summary) + 1,] = network_infected_summary

infected_summary
```

